Some examples in extended logic counting classes

### 2016-12-19 LICS/appendix.tex
19 dic

\textbf{Proof of Theorem \ref{eqso-fo-cap}}

\textbf{Proof of Theorem \ref{eqso-fo-cap-not}}

\textbf{Proof of Theorem \ref{eqso-eso-cap}}

\textbf{Proof of Theorem \ref{one-sigma-zero}}

>>>>

\subsection{Proof of Theorem \ref{eqso-fo-cap}}

For condition (1) in \ref{cap-def}, we will construct a recursive non-deterministic algorithm $M_{\alpha}$ for each $\eqso(\fo)$ formula $\alpha$ over a signature $\R$. This machine, for each $\A\in\ostr[\R]$, each first-order assignment $v$ over $\A$ and each second-order assignment $V$ over $\A$, on input $(\A,v,V)$ accepts in $\sem{\alpha}(\A,v,V)$ of its non-deterministic paths. Suppose $\A$ has domain $A$. If $\alpha$ is a $\fo$-formula $\varphi$, then the algorithm checks if $(\A,v,V)\models\varphi$ deterministically in polynomial time, and accepts if and only if it holds true. If $\alpha$ is a constant $s$, it produces $s$ branches and accepts in all of them. If $\alpha = (\beta \add \gamma)$, then it chooses between 0 or 1, if it is 0 (1), it simulates $M_{\beta}$ ($M_{\gamma}$) on input $(\A,v,V)$. If $\alpha = (\beta \mult \gamma)$, it simulates $M_{\beta}$ on input $(\A,v,V)$ and on each accepting path, it continues simulating $M_{\gamma}$ on input $(\A,v,V)$. If $\alpha = \sa{x}\beta$, it chooses $a\in A$ and simulates $M_{\beta}$ on input $(\A,v[a/x],V)$. If $\alpha = \pa{x}\beta$, it simulates $M_{\beta}$ on input $(\A,v[a/x],V)$ consecutively for each $a\in A$. If $\alpha = \sa{X}\beta$, it chooses $B\in A^{arity(X)}$ and simulates $M_{\beta}$ on input $(\A,v,V[B/X])$. This covers all possible cases for $\alpha$. Let $\alpha$ be a formula in $\eqso(\fo)$ over a signature $\R$ and let $f$ be a function over $\R$ such that $f(\enc(\A))$ is equal to the accepting paths of $M_{\alpha}$ on input $(\A,v,V)$ for some first-order assignment $v$ over $\A$ and some second-order assignment $V$ over $\A$. We have that $f$ is a $\shp$ function over $\R$ and that for every $\A\in\ostr[\R]$: $f(\enc(\A)) = \alpha(\enc(\A))$.

For condition (2) in \ref{cap-def}, note that Saluja et al. \cite{DBLP:journals/jcss/SalujaST95} proved that $\shp = \#\fo$. We also have that $\sqso(\fo)$ captures $\#\fo$ over ordered structures so for each $f\in \shp$ let $\alpha \in \sqso(\fo)$ be its corresponding formula. Since this formula is also in $\eqso(\fo)$, the condition holds.

\subsection{Proof of Theorem \ref{eqso-eso-cap}}

For condition (1) in \ref{cap-def}, we will construct a recursive non-deterministic algorithm $M_{\alpha}$ for each $\eqso(\eso)$ formula $\alpha$ over a signature $\R$. This machine, for each $\A\in\ostr[\R]$, each first-order assignment $v$ over $\A$ and each second-order assignment $V$ over $\A$ on input $(\A,v,V)$, accepts in $\sem{\alpha}(\A,v,V)$ of its non-deterministic outputs. Suppose $\A$ has domain $A$. On each of the steps the algorithm adds the current configuration to the output tape. If $\alpha$ is a $\eso$-formula $\varphi$ it checks if $(\A,v,V)\models\varphi$ non-deterministically in polynomial time without modifying the output tape, and accepts if and only if the condition holds true. If $\alpha$ is a constant $s$, then the algorithm produces $s$ branches and accepts in all of them. If $\alpha = (\beta \add \gamma)$, then it chooses between 0 or 1, if it is 0 (1), it simulates $M_{\beta}$ ($M_{\gamma}$) on input $(\A,v,V)$. If $\alpha = (\beta \mult \gamma)$, it simulates $M_{\beta}$ on input $(\A,v,V)$ and on each accepting path, it continues simulating $M_{\gamma}$ on input $(\A,v,V)$. If $\alpha = \sa{x}\beta$, it chooses $a\in A$ and simulates $M_{\beta}$ on input $(\A,v[a/x],V)$. If $\alpha = \pa{x}\beta$, it simulates $M_{\beta}$ on input $(\A,v[a/x],V)$ consecutively for each $a\in A$. If $\alpha = \sa{X}\beta$, it chooses $B\in A^{arity(X)}$ and simulates $M_{\beta}$ on input $(\A,v,V[B/X])$. This covers all possible cases for $\alpha$. Let $\alpha$ be a formula in $\eqso(\eso)$ over a signature $\R$ and let $f$ be a function over $\R$ such that $f(\enc(\A))$ is equal to the number of accepting outputs of $M_{\alpha}$ on input $(\A,v,V)$ for some first-order assignment $v$ over $\A$ and some second-order assignment $V$ over $\A$. We have that $f$ is a $\spp$ function over $\R$ and that for every $\A\in\ostr[\R]$: $f(\enc(\A)) = \alpha(\enc(\A))$.

For condition (2) in \ref{cap-def}, note that Compton and Graedel \cite{DBLP:journals/jcss/ComptonG96} proved that $\spp = \#\eso$. We also have that $\sqso(\eso)$ captures $\#\eso$ over ordered structures so for each $f\in \spp$ let $\alpha \in \sqso(\eso)$ be its corresponding formula. Since this formula is also in $\eqso(\eso)$, the condition holds.

\subsection{Proof of Theorem \ref{qso-pfp-cap}}

For condition (1) in \ref{cap-def}, we will construct a recursive non-deterministic algorithm $M_{\alpha}$ for each $\qso(\pfp)$ formula $\alpha$ over a signature $\R$. This machine, for each $\A\in\ostr[\R]$, each first-order assignment $v$ over $\A$ and each second-order assignment $V$ over $\A$, on input $(\A,v,V)$ accepts in $\sem{\alpha}(\A,v,V)$ of its non-deterministic paths. Suppose $\A$ has domain $A$. If $\alpha$ is a $\pfp$-formula $\varphi$, then the algorithm checks if $(\A,v,V)\models\varphi$ deterministically in polynomial space \cite{DBLP:conf/lics/AbiteboulV89}\martin{no estoy seguro de esta cita}, and accepts if and only if it holds true. If $\alpha$ is a constant $s$, it produces $s$ branches and accepts in all of them. If $\alpha = (\beta \add \gamma)$, then it chooses between 0 or 1, if it is 0 (1), it simulates $M_{\beta}$ ($M_{\gamma}$) on input $(\A,v,V)$. If $\alpha = (\beta \mult \gamma)$, it simulates $M_{\beta}$ on input $(\A,v,V)$ and on each accepting path, it continues simulating $M_{\gamma}$ on input $(\A,v,V)$. If $\alpha = \sa{x}\beta$, it chooses $a\in A$ and simulates $M_{\beta}$ on input $(\A,v[a/x],V)$. If $\alpha = \pa{x}\beta$, it simulates $M_{\beta}$ on input $(\A,v[a/x],V)$ consecutively for each $a\in A$. If $\alpha = \sa{X}\beta$, it chooses $B\in A^{arity(X)}$ and simulates $M_{\beta}$ on input $(\A,v,V[B/X])$. If $\alpha = \pa{X}\beta$, it simulates $M_{\beta}$ on input $(\A,v,V[B/X])$ consecutively for each $B\in A^{arity(X)}$. This covers all possible cases for $\alpha$, and each of these steps can be computed in polynomial space. Let $\alpha$ be a formula in $\qso(\pfp)$ over a signature $\R$ and let $f$ be a function over $\R$ such that $f(\enc(\A))$ is equal to the accepting paths of $M_{\alpha}$ on input $(\A,v,V)$ for some first-order assignment $v$ over $\A$ and some second-order assignment $V$ over $\A$. We have that $f$ is a $\shpspace$ function over $\R$, which implies that $f$ is also a $\fpspace$ function over $\R$ \cite{DBLP:journals/siamcomp/Ladner89}, and that for every $\A\in\ostr[\R]$: $f(\enc(\A)) = \alpha(\enc(\A))$.

\subsection{Proof of Theorem \ref{one-sigma-zero}}

### 2016-12-19 LICS/counting.tex
19 dic

Saluja et. al \cite{DBLP:journals/jcss/SalujaST95} define a family of counting classes $\#\LL$ for each fragment $\LL$ of $\fo$. For a formula $\varphi(x,X)$ over $\R$, the function $f_{\varphi(x,X)}$ is defined as
\[
f_{\varphi(x,X)}(\A) = \vert \{\langle e,P\rangle\mid \A\models\varphi(e,P)\}\vert.
\]
for each $\A\in\ostr[\R]$. A function $f:\Sigma^*\to\nat$ is in $\#\LL$ if there exists an $\LL$ formula $\varphi(x,X)$ such that $f = f_{\varphi(x,X)}$. Let $\sqso(\LL)$ be a fragment of $\qso$ defined by the following grammar:
\[
\alpha := \sa{x}\varphi \ \mid \ \sa{X} \varphi \ \mid \ \sa{x}\alpha \ \mid \ \sa{X}\alpha,
\]
where $\varphi$ is an $\LL$-formula. It can be seen that $\sqso(\LL)$ captures $\#\LL$ over ordered structures and that $\sqso(\LL) \subseteq \eqso(\LL)$.

@@

>>>>

Saluja et. al \cite{DBLP:journals/jcss/SalujaST95} define a family of counting classes $\#\LL$ for each fragment $\LL$ of $\so$. For a formula $\varphi(x,X)$ over $\R$, the function $f_{\varphi(x,X)}$ is defined as
\[
f_{\varphi(x,X)}(\A) = \vert \{\langle e,P\rangle\mid \A\models\varphi(e,P)\}\vert.
\]
for each $\A\in\ostr[\R]$. A function $f:\Sigma^*\to\nat$ is in $\#\LL$ if there exists an $\LL$ formula $\varphi(x,X)$ such that $f = f_{\varphi(x,X)}$. Let $\sqso(\LL)$ be a fragment of $\qso$ defined by the following grammar:
\[
\alpha := \sa{x}\varphi \ \mid \ \sa{X} \varphi \ \mid \ \sa{x}\alpha \ \mid \ \sa{X}\alpha,
\]
where $\varphi$ is an $\LL$-formula. It can be seen that $\sqso(\LL)$ captures $\#\LL$ over ordered structures and that $\sqso(\LL) \subseteq \eqso(\LL)$.

@@

\begin{corollary} \label{eqso-lfp-cap}
	$\eqso(\lfp)$ captures $\shp$ over ordered structures.
\end{corollary}
